---
title: "EIA categories"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EIA categories}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", out.width = "100%",
  message = FALSE, warning = FALSE, error = FALSE
)
library(eia)
options(eia_antidos = 0)
```

## Top level categories

It is helpful to be aware of various categories of data that are available. Each category has a unique ID number that is required to access associated data. The first call below does not include an ID. The result is a list of two data frames. The first is metadata associated with that position in the category hierarchy. The second is the child category information.

```{r cats0, eval=FALSE}
library(eia)
# eia_set_key("yourkey") # set your API key
```

Here is the top-level category information.

```{r cats1}
eia_cats()
```

## Child categories

The child category IDs can be used to query results from that category.

```{r cats3}
eia_cats(0)
```

View the immediate child categories for a given parent category.

```{r cats4}
eia_child_cats(389)
```

## Parent categories

View all parent categories for a given child category.

```{r cats5}
eia_parent_cats(742)
```

## Output format

The default is to return tidy data in a tibble data frame. For `eia_cats` you can set `tidy = FALSE` to return the list returned by `jsonlite::fromJSON` without any further processing or `tidy = NA` to return the raw JSON data as a character string.

```{r cats6}
eia_cats(tidy = FALSE)
cat(eia_cats(tidy = NA))
```
