---
title: "Introduction to eia"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to eia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", out.width = "100%",
  message = FALSE, warning = FALSE, error = FALSE
)
library(eia)
options(eia_antidos = 0)
```

This vignette provides a brief overview to the most important functions in `eia`. Other vignettes go into greater depth on specific topics and API endpoints.

## EIA categories

Pulling data from the US Energy Information Administration (EIA) API requires a registered API key. A key can be obtained at no cost [here](https://www.eia.gov/opendata/register.php).A valid email and agreement to the API Terms of Service is required to obtain a key.

Once you have your API key, it is helpful to be aware of various categories of data that are available. Each category has a unique ID number that is required to access associated data. The first call below does not include an ID. The result is a list of two data frames. The first is metadata associated with that position in the category hierarchy. The second is the child category information.

Remember to store your API key somewhere secure. Do not commit it to a repository or otherwise share it. For example, you could store it in your `.Renviron` file.

```{r cats0}
library(eia)
key <- Sys.getenv("EIA_KEY")
```

Here is the top-level category information:

```{r cats1}
eia_cats(key)
```

The child category IDs can be used to query results from that category.

```{r cats2}
eia_cats(key, 0)
```

## EIA time series data

Add later...

## EIA geosets

Add later...
