---
title: "EIA geosets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EIA geosets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", out.width = "100%",
  message = FALSE, warning = FALSE, error = FALSE
)
library(tibble)
library(eia)
options(eia_antidos = 0)
```

## Time series by region

Geosets are metadata structures relating series together. Requesting series using the `geoset` API endpoint is much the same as requesting series data using the `series` endpoint. Instead of `eia_series`, use `eia_geoset`.

The main difference is that you must provide a geoset ID to `id` and a `region` argument. Both may be vectors of multiple series and regions. The function returns the combination of time series datasets and regions that exist. The API will not return all geographic entities associated with a geoset. You are required to specify which region(s) you want and they must be associated with the given geoset ID.

```{r geoset}
library(eia)
# eia_set_key("yourkey") # set your API key
id <- "ELEC.GEN.ALL-99.A"
region <- c("USA-CA", "USA-NY")
(x <- eia_geoset(id[1], region[1], n = 3))
```

## Groups of regions

If you want data for all fifty states for example, you can set `region = "USA"`; you do not need to make a vector of all fifty state IDs. However, if you want certain region subsets, there are other options besides making a vector of these values. For example, the `eia_geoset` accepts shorthand descriptions of specific, popular subsets of states; so popular in fact that their labels ship with R itself in the `datasts` package.

You can provide simple state abbreviations (without the `USA-` prefix), state names, and more to the point, US Census regions and divisions. These are two hierarchical sets of US states.

```{r census}
tibble(state.abb, state.name, state.region, state.division)
```

Provide the associated label and `eia_geoset` recognizes subsets of US states. Even this can be a vector. In the example below, `region` consists of Alaska plus the states belonging to the New England census division.

```{r geoset2}
(x <- eia_geoset(id, c("AK", "New England"), n = 1))
x$data[[1]]

region <- "Middle Atlantic"
x <- eia_geoset(id, region, n = 12)
select(x, region, data) %>% unnest()

unnest(x, cols = data) %>% 
  ggplot(aes(date, value, color = region)) +
  geom_line() + 
  labs(y = d$units[1], title = "Net electricity generation, all fuels")
```

## Relations

There is also a `relation` argument that accepts an optional relation ID. If one is provided, `eia_geoset` will switch to the API `relation` endpoint. A relation is another metadata structure that applies to geosets and relates summary statistics associated with geoset IDs to their composite statistics. This makes it easier to obtain variables that facet the data, e.g., by sector or fuel type.

The EIA `relation` API endpoint is officially supported according to the online EIA API documentation, but unfortunately that endpoint does not appear to function at the time of current package release.
